News Across Languages - Cross-Lingual Document Similarity and Event Tracking

### 2015-03-24 rupnik2015.tex
- changes througout the document

Although there are a number of services that aggregate news by identifying clusters of similar articles, there are almost no services that provide linking of clusters over different languages. Google News as well as Yahoo! News are able to identify clusters of articles about same event, but they offer no linking of clusters across languages. The only service that we found, which provides cross-lingual cluster linking, is the European Media Monitor (EMM)~\cite{Steinberger2008}. EMM clusters articles in 60 languages and then tries to determine which clusters of articles in different languages describe the same event. To achieve cluster linking, EMM uses three different language independent vector representations for each cluster. First vector contains the weighted list of references to countries mentioned in the articles, while the second vector contains the weighted list of mentioned people and organizations. The last vector contains the weighted list of Eurovoc subject domain descriptors. These descriptors are topics, such as \emph{air transport}, \emph{EC agreement}, \emph{competition} and \emph{pollution control} into which articles are automatically categorized~\shortcite{Pouliquen2006}. Similarity between clusters is then computed using a linear combination of the cosine similarities computed on the lists. If the similarity is above the threshold, the clusters are linked. Compared to EMM, our approach uses document similarities to obtain a small set of potentially equivalent clusters. Additionally, we do not decide if two clusters are equivalent based on a hand-set threshold on a similarity value -- instead we use a classification model that uses a larger set of features related to the tested clusters.

@@

The second stage of the algorithm (described in the Algorithm~\ref{cluster_merge_algo2}.) determines which (if any) of the candidate clusters are equivalent to $c_i$. We approach this as a supervised learning problem. For each candidate cluster $c_j \in C'$, we compute a vector of learning features that should be indicative of whether the $c_i$ and $c_j$ are equivalent or not. The details of our chosen features will be described in the following Section. Once these features are computed, we apply a classification model which was trained using a  dataset  of hand-labeled examples. If the model predicts that tested clusters are equivalent, the $c_j$ is added to the set $C'$.

@@

The dataset provided by the annotators contains 808 examples, of which 402 are equivalent clusters pairs and 406 are not. Clusters in each learning example are either in English, Spanish or German. Although Event Registry imports articles in other languages as well, we restricted our experiments to these three languages, since they have the highest volume of articles which makes the training problem similarly hard \todo{doesnt this make it easier?}.

@@

\item {\bf Concept related features}. As mentioned before, articles that are imported into Event Registry are first semantically annotated by linking mentioned entities and keywords to the corresponding Wikipedia pages. Whenever Barack Obama is, for example, mentioned in the article, the article is annotated with a link to his Wikipedia page (\emph{http://en.wikipedia.org/wiki/Barack\_Obama}). In the same way all mentions of people, locations, organizations and even ordinary keywords (e.g. bank, tax, ebola, plane, company) are annotated. Although the Spanish article about Obama will be annotated with his Spanish version of the Wikipedia page, in many cases we can link the Wikipedia pages to their English versions. This can be done since Wikipedia itself provides information regarding which pages in different languages represent the same concept/entity. Using this approach, the word ``avi\'on'' in a Spanish article will be annotated with the same concept as the word ``plane'' in an English article. Although the articles are in different languages, the annotations can therefore provide a shared vocabulary that can be used to compare articles/clusters. By analyzing all the articles in clusters $c_i$ and $c_j$, we can identify the most relevant entities and keywords for each cluster. Additionally, we can also assign weights to the concepts based on how frequently they occur in the articles in the cluster. From the list of relevant concepts and corresponding weights, we consider the following features: \texttt{entityCosSim} which represents the cosine similarity on the vector of entities mentioned in both clusters; \texttt{keywordCosSim} which represents the cosine similarity on the keywords only; for cases where the weights are not relevant, we have also added two additional features - (\texttt{entityJaccardSim} and \texttt{keywordJaccardSim}), where Jaccard similarity is used instead of the cosine similarity.

\item {\bf Miscellaneous features}. This group contains three miscellaneous features that seem discriminative but are unrelated to the previous two groups. The first feature is related to the event location. By considering the locations mentioned in the articles in a cluster, we are able to estimate where an event occurred. The \texttt{hasSameLocation} feature is a boolean variable that is true when the location of the event in both clusters is the same. Another important feature is related to the time when the articles in the clusters were published. For each cluster, we first compute the average publication time and date of the articles. The \texttt{timeDiff} is then defined as the absolute difference in hours between the two computed dates. The final feature we consider is computed by analyzing the mentions of dates in the articles. Using an extensive set of regular expressions, we are able to detect mentions of dates in different forms. To compute the feature \texttt{sharedDates} we start by identifying the list of dates mentioned in articles in each cluster separately. The value of \texttt{sharedDates} is then determined as the ratio of all dates that are detected in both clusters.\todo{this should be made more explicit/}

@@

\subsection{Experiments With Missing Alignment Data}
%\emph{Andrej started working on this, done by Monday}
%{\oldText{Taken from XLite paper}
%To investigate the empirical performance of our algorithm we will test it on a large-scale, real-world multilingual dataset that we extracted from Wikipedia  by  using so called 'inter-language' links as an alignment.
 In this subsection we will investigate the empirical performance of cca hub approach. We will demonstrate that this approach can be successfully applied even in the case of missing alignment information.
 To this purpose we select a subset of Wikipedia languages containing three major languages, English--\emph{en} (hub language), Spanish--\emph{es}, Russian--\emph{ru}, and five minority (in the sense of Wikipedia sizes) languages, Slovenian--\emph{sl}, Piedmontese--\emph{pms}, Waray-Waray--\emph{war} (all with about 2 million native speakers), Creole--\emph{ht} (8 million native speakers), and Hindi--\emph{hi} (180 million native speakers). For preprocessing we remove the documents that contain less than 20 different words (stubs) and remove words occurring in less than 50 documents as well as the top 100 most frequent words (in each language separately). We represent the documents as normalized TFIDF\cite{Salton88term-weightingapproaches} weighted vectors.
 %These are languages with Wikipedia article sizes comparable to the minority languages.  (since the number of speakers does not directly correlate with the number of Wikipedia articles). The prime hub candidate is the English %language which is well aligned with all other Wikipedia languages, although the alignment quality varies quite a bit. Furthermore, we remove the documents that contain less than 20 different words and remove words that are too %infrequent as well as the top 100 most frequent words in the vocabularies.
 Although the English language is well aligned with all Wikipedia languages, we must note that quality of alignment varies quite a bit, especially in the case of small Wikipedias.
 Furthermore, we call the document consisting of less than 20 different words, a stub. This documents are typically garbage, the titles of the columns in the table, remains of the parsing process, or Wikipedia articles with very little or no information contained in one or two sentences.

@@

In Section~\ref{sec:models} we have described three main algorithms for identifying similar articles in different languages. These algorithms were $k$-means, LSI and CCA. To test which of these algorithms performs best we performed the following test. Using each algorithm, we analyzed all articles in Event Registry to find the most similar articles in other languages. The values of the cross-lingual article linking features (\texttt{linkCount} and \texttt{avgSimScore}) were then recomputed for all annotated learning examples. A dataset with only these two features and the label was constructed and used to test how accurate prediction models can be induced using the data. The results of the experiment are shown in Table~\ref{table:linkingEvalAlgos}. The table shows for each of the algorithms the obtained classification accuracy, precision and recall. We can see that the most informative features for cluster linking are obtained using the CCA algorithm, while $k$-means performs the worse. Firstly we verified the AMRR (average of mean reciprocal ranks over all language pairs) performance of approaches by holding out $15000$ aligned test documents and using $300000$ aligned documents
 as the training set. Figure~\ref{pc:AMRR} shows AMRR score as the function of the number of feature vectors. It is clear that CCA outperforms LSI approach and K-means lags far behind.
Furthermore we inspected how number of features influences the accuracy of cluster linking. As we can see from Table~\ref{table:linkingEvalAlgos} choosing number of features vectors larger than $500$ barely effects linking performance. That is probably due to overfitting. Moreover we tested the performance of detecting large/small cluster linking, see Table~\ref{table:linkingEvalAlgosLargeSmall}.  As expected we obtained that it is harder to link small clusters and we need better retrieval. Hub CCA outperforms other methods in this sense also, probably as it has the best retrieval performance, see Figure~\ref{pic:AMRR}.

@@

\begin{table}[h]
\caption{Accuracy of large/small cluster linking with $500$ features obtained from different cross-lingual similarity algorithms}
\label{table:linkingEvalAlgosLargeSmall}
\begin{center}
\begin{tabular}{|c|c|c|c|}
  \hline
  \cline{1-4}
  Models & CA & Precision & Recall \\ \cline{1-4}
  CCA      & 81.2 - 77.8 \% & 80.5 - 74.5 \%  & 91.3 - 57.5 \%  \\ \cline{1-4}
  LSI      & 82.8 - 76.4 \% & 81.3 - 70.9 \%  & 93.1 - 57.5 \%  \\ \cline{1-4}
 $k$-means & 75.5 - 71.2 \% & 72.8 - 70.8 \%  & 95.3 - 36.2 \%  \\ \cline{1-4}
\end{tabular}
\end{center}
\end{table}

@@

\item \textbf{Features} - The experiments show that the hub CCA based features present good baseline, which can greatly benefit from additional semantic extraction based features. Even though the CCA based features do not greatly improve the system performance when the semantic features are used, there are two main benefits in the approach: the linking process can be sped up (smaller candidate cluster sets), and in the case when semantic extraction is not possible.

\item \textbf{Future work} - Currently the system is loosely coupled - the language component is built independently of the rest of the system, in particular the linking component. It is possible that better embeddings can be obtained by methods that jointly optimize a classification task and the embedding.

    Another point of interest is to evaluate the system on languages with scarce linguistic resources, where semantic annotation might not be available. For this reason, the labelled dataset of linked clusters should to be extended first. The mate retrieval evaluation showed that even for language pairs with no training set overlap, the hub cca recovers some signal.

>>>>

Although there are a number of services that aggregate news by identifying clusters of similar articles, there are almost no services that provide linking of clusters over different languages. Google News as well as Yahoo! News are able to identify clusters of articles about same event, but they offer no linking of clusters across languages. The only service that we found, which provides cross-lingual cluster linking, is the European Media Monitor (EMM)~\cite{Steinberger2008}. EMM clusters articles in 60 languages and then tries to determine which clusters of articles in different languages describe the same event. To achieve cluster linking, EMM uses three different language independent vector representations for each cluster. First vector contains the weighted list of references to countries mentioned in the articles, while the second vector contains the weighted list of mentioned people and organizations. The last vector contains the weighted list of Eurovoc subject domain descriptors. These descriptors are topics, such as \emph{air transport}, \emph{EC agreement}, \emph{competition} and \emph{pollution control} into which articles are automatically categorized~\shortcite{Pouliquen2006}. Similarity between clusters is then computed using a linear combination of the cosine similarities computed on the lists. If the similarity is above the threshold, the clusters are linked. Compared to EMM, our approach uses document similarities to obtain a small set of potentially equivalent clusters. Additionally, we do not decide if two clusters are equivalent based on a hand-set threshold on a similarity value -- instead we use a classification model that uses a larger set of features related to the tested pair of clusters.

@@

The second stage of the algorithm (described in the Algorithm~\ref{cluster_merge_algo2}.) determines which (if any) of the candidate clusters are equivalent to $c_i$. We treat this task as a supervised learning problem. For each candidate cluster $c_j \in C'$, we compute a vector of learning features that should be indicative of whether the $c_i$ and $c_j$ are equivalent or not. The details of our chosen features will be described in the following section. Once these features are computed, we apply a classification model which was trained using a  dataset  of hand-labeled examples. If the model predicts that tested clusters are equivalent, the $c_j$ is added to the set $C'$.

@@

The dataset provided by the annotators contains 808 examples, of which 402 are equivalent clusters pairs and 406 are not. Clusters in each learning example are either in English, Spanish or German. Although Event Registry imports articles in other languages as well, we restricted our experiments to these three languages. We chose only these three languages since they have very large number of articles and clusters per day which makes the cluster linking problem hard due to large number of possible links.

@@

\item {\bf Concept related features}. Articles that are imported into Event Registry are first semantically annotated by linking mentioned entities and keywords to the corresponding Wikipedia pages. Whenever Barack Obama is, for example, mentioned in the article, the article is annotated with a link to his Wikipedia page (\emph{http://en.wikipedia.org/wiki/Barack\_Obama}). In the same way all mentions of people, locations, organizations and even ordinary keywords (e.g. bank, tax, ebola, plane, company) are annotated. Although the Spanish article about Obama will be annotated with his Spanish version of the Wikipedia page, in many cases we can link the Wikipedia pages to their English versions. This can be done since Wikipedia itself provides information regarding which pages in different languages represent the same concept/entity. Using this approach, the word ``avi\'on'' in a Spanish article will be annotated with the same concept as the word ``plane'' in an English article. Although the articles are in different languages, the annotations can therefore provide a language-independent vocabulary that can be used to compare articles/clusters. By analyzing all the articles in clusters $c_i$ and $c_j$, we can identify the most relevant entities and keywords for each cluster. Additionally, we can also assign weights to the concepts based on how frequently they occur in the articles in the cluster. From the list of relevant concepts and corresponding weights, we consider the following features: \texttt{entityCosSim} which represents the cosine similarity on the vector of entities mentioned in both clusters; \texttt{keywordCosSim} which represents the cosine similarity on the keywords only; for cases where the weights are not relevant, we have also added two additional features - (\texttt{entityJaccardSim} and \texttt{keywordJaccardSim}), where Jaccard similarity is used instead of the cosine similarity.

\item {\bf Miscellaneous features}. This group contains three miscellaneous features that seem discriminative but are unrelated to the previous two groups. The first feature is related to the event location. By considering the locations mentioned in the articles in a cluster, we are able to estimate where an event occurred. The \texttt{hasSameLocation} feature is a boolean variable that is true when the location of the event in both clusters is the same. Another important feature is related to the time when the articles in the clusters were published. For each cluster, we first compute the average publication time and date of the articles. The \texttt{timeDiff} is then defined as the absolute difference in hours between the two computed dates. The final feature we consider is computed by analyzing the mentions of dates in the articles. Using an extensive set of regular expressions, we are able to detect mentions of dates in different forms. To compute the feature \texttt{sharedDates} we start by identifying the list of dates mentioned in articles in each cluster separately. The value of \texttt{sharedDates} is then determined as the Jaccard similarity on the two lists of dates.

@@

\subsection{Experiments With Missing Alignment Data}
%\emph{Andrej started working on this, done by Monday}
%{\oldText{Taken from XLite paper}
%To investigate the empirical performance of our algorithm we will test it on a large-scale, real-world multilingual dataset that we extracted from Wikipedia  by  using so called 'inter-language' links as an alignment.
 In this subsection we will investigate the empirical performance of CCA hub approach. We will demonstrate that this approach can be successfully applied even in the case of missing alignment information.
 To this purpose we select a subset of Wikipedia languages containing three major languages, English--\emph{en} (hub language), Spanish--\emph{es}, Russian--\emph{ru}, and five minority (in the sense of Wikipedia sizes) languages, Slovenian--\emph{sl}, Piedmontese--\emph{pms}, Waray-Waray--\emph{war} (all with about 2 million native speakers), Creole--\emph{ht} (8 million native speakers), and Hindi--\emph{hi} (180 million native speakers). For preprocessing we remove the documents that contain less than 20 different words (stubs) and remove words occurring in less than 50 documents as well as the top 100 most frequent words (in each language separately). We represent the documents as normalized TFIDF\cite{Salton88term-weightingapproaches} weighted vectors.
 %These are languages with Wikipedia article sizes comparable to the minority languages.  (since the number of speakers does not directly correlate with the number of Wikipedia articles). The prime hub candidate is the English %language which is well aligned with all other Wikipedia languages, although the alignment quality varies quite a bit. Furthermore, we remove the documents that contain less than 20 different words and remove words that are too %infrequent as well as the top 100 most frequent words in the vocabularies.
 Although the English language is well aligned with all Wikipedia languages, we must note that quality of alignment varies quite a bit, especially in the case of small Wikipedias.
 Furthermore, we call the document consisting of less than 20 different words, a stub. This documents are typically garbage, the titles of the columns in the table, remains of the parsing process, or Wikipedia articles with very little or no information contained in one or two sentences.

@@

In Section~\ref{sec:models} we have described three main algorithms for identifying similar articles in different languages. These algorithms were $k$-means, LSI and CCA. To test which of these algorithms performs best we performed the following test. Using each algorithm, we analyzed all articles in Event Registry to find the most similar articles in other languages. The values of the cross-lingual article linking features (\texttt{linkCount} and \texttt{avgSimScore}) were therefore recomputed for all annotated learning examples using all three methods. A dataset with only these two features and the label was constructed and used to test how accurate prediction models can be induced using the data. The results of the experiment are shown in Table~\ref{table:linkingEvalAlgos}. The table shows for each of the algorithms the obtained classification accuracy, precision and recall. We can see that the most informative features for cluster linking are obtained using the CCA algorithm, while $k$-means performs the worse. Firstly we verified the AMRR (average of mean reciprocal ranks over all language pairs) performance of approaches by holding out $15000$ aligned test documents and using $300000$ aligned documents as the training set. Figure~\ref{pc:AMRR} shows AMRR score as the function of the number of feature vectors. It is clear that CCA outperforms LSI approach and K-means lags far behind.\todo{the sentences after "Firstly we verified" are not clear - seem out of context}

Furthermore we inspected how number of features influences the accuracy of cluster linking. As we can see from Table~\ref{table:linkingEvalAlgos} choosing number of features vectors larger than $500$ barely effects linking performance. That is probably due to overfitting.

Additionally we also tested how accurately can we link clusters depending on the cluster size. For this experiment we separated the learning examples into two datasets -- one dataset contained clusters where the sum of articles in both clusters is below 20 articles (356 learning examples) while the other dataset contained the rest of the examples (452 examples). 10 fold cross validation was then done separately for each of the two datasets. Results are shown in Table~\ref{table:linkingEvalAlgosLargeSmall}.  As expected, we can see that it is harder to link small clusters than large clusters since we have less articles that provide the necessary information. We can also see that CCA and LSI outperform the $k$-means approach, which is likely due to their high retrieval performance (see Figure~\ref{pic:AMRR}).

@@

\begin{table}[h]
\caption{Accuracy of cluster linking using $500$ features on two datasets containing large (left number) and small (right number) clusters}
\label{table:linkingEvalAlgosLargeSmall}
\begin{center}
\begin{tabular}{|c|c|c|c|}
  \hline
  \cline{1-4}
  Models & CA & Precision & Recall \\ \cline{1-4}
  CCA      & 81.2 - 77.8 \% & 80.5 - 74.5 \%  & 91.3 - 57.5 \%  \\ \cline{1-4}
  LSI      & 82.8 - 76.4 \% & 81.3 - 70.9 \%  & 93.1 - 57.5 \%  \\ \cline{1-4}
 $k$-means & 75.5 - 71.2 \% & 72.8 - 70.8 \%  & 95.3 - 36.2 \%  \\ \cline{1-4}
\end{tabular}
\end{center}
\end{table}

@@

\item \textbf{Features} - The experiments show that the hub CCA based features present good baseline, which can greatly benefit from additional semantic extraction based features. Even though the CCA based features do not greatly improve the system performance when the semantic features are used, there are two main benefits of the approach. By obtaining a small set of candidate clusters, the cluster linking task can be performed significantly compared to exhaustive evaluation of all cluster pairs. Additionally, for many languages, the named entity and keyword extraction is simply not available which prevents us from using them as learning features.

\item \textbf{Future work} - Currently the system is loosely coupled -- the language component is built independently of the rest of the system, in particular the linking component. It is possible that better embeddings can be obtained by methods that jointly optimize a classification task and the embedding.

    Another point of interest is to evaluate the system on languages with scarce linguistic resources, where semantic annotation might not be available. For this task, the labeled dataset of linked clusters should be extended first. The mate retrieval evaluation showed that even for language pairs with no training set overlap, the hub CCA recovers some signal.
